### 2021-04-08 02_abstract/abstract.tex
Spell- and grammer-checked chapter 1

In a first step, the climate sensitivity (\ie{} the temperature response of the
climate system to an external forcing) is evaluated in the latest generation of
\acsp{ESM} from \acs{CMIP}6. For the \acfAbstract{ECS}, which is defined as the
equilibrium temperature response that follows a doubling of the atmospheric
\acfAbstract{CO2} concentration, a \acfAbstract{MMM} of $3.74 \unit{K}$ and a
multi-model range of $\rangeunit{1.8}{5.6}{K}$ are found. These values are
higher than in any previous \acs{CMIP} ensemble before. Moreover, a third of
the analyzed \acs{CMIP}6 models exceed the upper bound of the likely \acs{ECS}
range of $\rangeunit{1.5}{4.5}{K}$ assessed by the \acfAbstract{IPCC} Fifth
\acl{AR} (\acs{AR}5) from 2013. Similarly, the transient response of the
climate system to a doubling of \acs{CO2}, also known as \acfAbstract{TCR},
shows an inter-model range of $\rangeunit{1.3}{3.0}{K}$ with an upper bound
again higher than the likely range assessed in \acs{AR}5 of
$\rangeunit{1.0}{2.5}{K}$. Possible reasons for the increased climate
sensitivity in many \acs{CMIP}6 models are the addition of new prognostic
aerosol schemes that include aerosol-cloud interactions and changes in the
microphysical representation of mixed-phase clouds. The changes in the
mixed-cloud representation reduce the strong negative shortwave cloud phase
change feedback over the Southern Ocean that is present in climate models from
previous \acs{CMIP} generations.

@@

To overcome these issues of single-process-oriented emergent constraints, an
alternative approach based on \acfAbstract{ML} is introduced. Since this new
technique relies on a large number of data points to train the \acs{ML}
algorithm, the scalar climate sensitivity expressed as \acs{ECS} or \acs{TCR}
is not an appropriate target variable. Therefore, \acfAbstract{GPP} as a
process that contributes to climate sensitivity is studied as an alternative.
\acs{GPP} is the largest flux of the terrestrial carbon uptake and slows down
global warming by removing \acs{CO2} from the atmosphere. In this analysis, an
existing emergent constraint on \acs{CO2} fertilization is combined with an
\acs{ML} approach to constrain the spatial variations of multi-model \acs{GPP}
projections. In the first step of the two-step approach, observed changes in
the \acs{CO2} seasonal cycle at \acl{KUM} are used to constrain the global mean
\acs{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in \acl{RCP}
8.5 simulations with \acsp{ESM} participating in \acs{CMIP}5 to
$\pmrangeunit{171}{12}{\GtCyr{}}$, compared to the unconstrained model range
$\rangeunit{156}{247}{\GtCyr{}}$. In a second step, an \acs{ML} model is used
to constrain gridded future absolute \acs{GPP} and gridded fractional \acs{GPP}
change in two independent approaches. For this, observational data is fed into
the \acs{ML} algorithm that has been trained on \acs{CMIP}5 data to learn
relationships between present-day physically relevant diagnostics and the
target variable. In a leave-one-model-out \acl{CV} approach, the \acs{ML} model
shows superior performance to the \acs{CMIP}5 \acs{MMM}. The new approach
predicts a higher \acs{GPP} increase in high latitudes and a lower \acs{GPP}
increase in regions closer to the equator.

>>>>

In a first step, climate sensitivity (\ie{} the temperature response of the
climate system to an external forcing) is evaluated in the latest generation of
\acsp{ESM} from \acs{CMIP}6. For the \acfAbstract{ECS}, which is an estimation
of the equilibrium temperature response that follows a doubling of the
atmospheric \acfAbstract{CO2} concentration, a \acfAbstract{MMM} of $3.74
\unit{K}$ and a multi-model range of $\rangeunit{1.8}{5.6}{K}$ are found. These
values are higher than in any previous \acs{CMIP} ensemble before. Moreover, a
third of the analyzed \acs{CMIP}6 models exceed the upper bound of the likely
\acs{ECS} range of $\rangeunit{1.5}{4.5}{K}$ assessed by the \acfAbstract{IPCC}
Fifth \acl{AR} (\acs{AR}5) from 2013. Similarly, the transient response of the
climate system to a doubling of \acs{CO2}, also known as \acfAbstract{TCR},
shows an inter-model range of $\rangeunit{1.3}{3.0}{K}$ with an upper bound
again higher than the likely range assessed in \acs{AR}5 of
$\rangeunit{1.0}{2.5}{K}$. Possible reasons for increased climate sensitivity
in many \acs{CMIP}6 models are the addition of new prognostic aerosol schemes
that include aerosol-cloud interactions and changes in the microphysical
representation of mixed-phase clouds. The changes in the mixed-cloud
representation reduce the strong negative shortwave cloud phase change feedback
over the Southern Ocean that is present in climate models from previous
\acs{CMIP} generations.

@@

To overcome these issues of single-process-oriented emergent constraints, an
alternative approach based on \acfAbstract{ML} is introduced. Since this new
technique relies on a large number of data points to train the \acs{ML}
algorithm, scalar climate sensitivity expressed as \acs{ECS} or \acs{TCR} is
not an appropriate target variable. Therefore, \acfAbstract{GPP} as a process
that contributes to climate sensitivity is studied as an alternative. \acs{GPP}
is the largest flux of the terrestrial carbon uptake and slows down global
warming by removing \acs{CO2} from the atmosphere. In this analysis, an
existing emergent constraint on \acs{CO2} fertilization is combined with an
\acs{ML} approach to constrain the spatial variations of multi-model \acs{GPP}
projections. In the first step of the two-step approach, observed changes in
the \acs{CO2} seasonal cycle at \acl{KUM} are used to constrain the global mean
\acs{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in \acl{RCP}
8.5 simulations with \acsp{ESM} participating in \acs{CMIP}5 to
$\pmrangeunit{171}{12}{\GtCyr{}}$, compared to the unconstrained model range
$\rangeunit{156}{247}{\GtCyr{}}$. In a second step, an \acs{ML} model is used
to constrain gridded future absolute \acs{GPP} and gridded fractional \acs{GPP}
change in two independent approaches. For this, observational data is fed into
the \acs{ML} algorithm that has been trained on \acs{CMIP}5 data to learn
relationships between present-day physically relevant diagnostics and the
target variable. In a leave-one-model-out \acl{CV} approach, the \acs{ML} model
shows superior performance to the \acs{CMIP}5 \acs{MMM}. The new approach
predicts a higher \acs{GPP} increase in high latitudes and a lower \acs{GPP}
increase in regions closer to the equator.

### 2021-04-08 ch01_introduction/introduction.tex
Spell- and grammer-checked chapter 1

Climate change is one of the greatest challenges for humankind today. The
warming of the climate system is \enquote{unequivocal}, and \enquote{many of
  the observed changes are unprecedented over decades to millennia}
\autocite{IPCC2014}. The changing climate increases the \enquote{likelihood of
  severe, pervasive and irreversible impacts for people, species and
  ecosystems} with \enquote{mostly negative impacts for biodiversity, ecosystem
services and economic development}, and amplifies \enquote{risks for
livelihoods and for food and human security} \autocite{IPCC2014}. Potential
drivers for climate change are all natural and anthropogenic substances and
processes that may alter the Earth's energy budget. Human influence on the
climate system is clear: Ever increasing emissions of \acp{GHG} since the end
of the pre-industrial era largely driven by economic and population growth
led to atmospheric concentrations of \ac{CO2}, \ac{CH4}, and \ac{N2O} that
are \enquote{unprecedented in at least the last 800000 years}
\autocite{IPCC2014}. The effects of these \ac{GHG} emissions and other
anthropogenic drivers have been \enquote{detected throughout the climate
system and are extremely likely to have been the dominant cause of the
observed warming since the mid-\nth{20} century} \autocite{IPCC2014}. The
greenhouse effect is based on the optical properties of the \acp{GHG}: While
the atmosphere is mostly transparent for the incoming solar (shortwave)
radiation, the outgoing infrared (longwave) radiation that is reflected from
the planet's surface is partly absorbed by the corresponding \ac{GHG}
molecules through an excitation of their corresponding vibrational modes and
re-emitted isotropically. This process traps energy near the surface and
leads to a warming of the Earth's surface and the lower atmosphere. The
individual impact of the different drivers of climate change is measured with
the so-called \emph{radiative forcing}, which quantifies the change in energy
fluxes caused by changes in these drivers relative to pre-industrial
conditions \autocite{IPCC2013}. The total radiative forcing is positive
(which corresponds to a warming) and its largest contribution is caused by
the increase in the atmospheric concentration of \ac{CO2} due to anthropogenic
fossil fuel emissions since the mid-\nth{18} century \autocite{IPCC2014}. Apart
from their physical warming effect on the climate, carbon-based \acp{GHG} like
\ac{CO2} also directly influence the global carbon cycle, an important
biogeochemical cycle of the Earth. A crucial flux of the carbon cycle is the
\emph{\ac{GPP}}, which describes the carbon uptake of the terrestrial biosphere
due to photosynthesis. Since this land carbon uptake absorbs about $30
\unit{\%}$ of the anthropogenic \ac{CO2} emissions in today's climate
\autocite{Friedlingstein2020}, this process substantially slows down global
warming and directly contributes to the magnitude of climate sensitivity. Other
important anthropogenic drivers of climate change are the emission of aerosols,
and land use/land cover changes. Apart from anthropogenic drivers, there are
also natural processes which impact the climate system like changes in the
solar activity or volcanic eruptions. However, there is clear evidence that
these natural drivers alone cannot explain the observed climate change
\autocite{Haustein2017}.

@@

Simulations that extrapolate the state of the climate system into the future
are called \emph{climate model projections}. These include idealized
simulations with only a prescribed change in the atmospheric \ac{CO2}
concentration (\eg{} an instantaneous doubling of \ac{CO2} or \iac{CO2}
increase of $1 \unit{\%}$ per year) as well as more realistic projections that
consider different future scenarios (\eg{} a fossil fuel--based future or a
scenario that is based on a sustainable development). Multi-model projections
from \ac{CMIP} show a large inter-model range in many variables that are
relevant for climate change \autocite{Collins2013, Flato2013}. A crucial and
policy-relevant example for this is the climate sensitivity, which refers to
the change in the \ac{GSAT} that results from a change in the radiative
forcing. Common metrics for this are the \emph{\ac{ECS}}, which describes the
equilibrium response of the climate system after a doubling of the atmospheric
\ac{CO2} concentration, and the \emph{\ac{TCR}}, which describes the transient
response of the system to \iac{CO2} doubling. In \acs{AR}5, the assessed range
of both quantities is large with $\rangeunit{1.5}{4.5}{K}$ for \ac{ECS} and
$\rangeunit{1.0}{2.5}{K}$ for \ac{TCR} \autocite{Stocker2013}. The
corresponding inter-model ranges from the \acs{CMIP}5 models show similar
results \autocite{Flato2013}. For this reason, a careful statistical evaluation
and further refinement of the output of multi-model climate projections is
necessary to reduce associated uncertainties. A state-of-the-art technique for
this is the \emph{emergent constraints} method, which uses a physically based
inter-model relationship between an observable quantity of the Earth system and
a target variable to reduce uncertainties in the target variable with
observations \autocite{Allen2002}. An alternative approach is the weighting of
climate models based on their performance (\ie{} the distance of one model to
observational products) and interdependence (\ie{} the distance of one model to
other climate models) \autocite{Knutti2017a}. These techniques form the
baseline for the new analyses and results presented in this thesis, which
partly utilize methods from a new emerging research field in climate sciences:
\ac{AI} and \ac{ML}.

@@

The aim of this thesis is to reduce uncertainties in multi-model climate
projections with observations by addressing the following three
\namecrefs{enum:01:question_1}:

@@

Parts of this thesis are already published in multiple peer-reviewed
publications (two first-author studies and six co-author studies). A complete
list of these publications is given on \cpageref{ch:bib:own}. Wherever material
from these studies is presented in this thesis, the pronoun \enquote{we} is
used to increase readability by avoiding the passive voice and to acknowledge
all involved contributors. However, unless stated otherwise, all content from
these publications (text, figures, and tables) shown in this thesis originate
from the author of this thesis. A detailed list of contributions to these
studies is given in the corresponding \namecrefs{ch:03:esmvaltool}.

This thesis is structured as follows: \Cref{ch:02:scientific_background}
introduces the scientific background. This includes relevant literature that is
used as a baseline for this thesis. \Cref{ch:03:esmvaltool} gives an overview
over the contributions made to the \ac{ESMValTool}, an open-source software for
the analysis of \acp{ESM}. These contributions helped improving the routine
evaluation of \acp{ESM} which is useful for the entire scientific community and
led to co-authorship in four peer-reviewed studies \autocite{Eyring2020,
  Lauer2020, Righi2020, Weigel2020}. \Cref{ch:04:papers_ecs_tcr_assessment}
covers the assessment of climate sensitivity metrics like the \ac{ECS} or
\ac{TCR} in the latest generation of \acp{ESM} from \acs{CMIP}6. This work is
already published in two scientific publications \autocite{Bock2020,
  Meehl2020}. Since the \ac{ECS} and \ac{TCR} are considerably higher in this
new climate model generation, \cref{ch:05:paper_ecs} describes the assessment
of emergent constraints on the \ac{ECS} for these \acs{CMIP}6 models and
compares these to results derived from \acs{CMIP}5 models. The content of this
\namecref{ch:05:paper_ecs} is published in \emph{Earth System Dynamics}
\autocite{Schlund2020a}. \Cref{ch:06:paper_gpp} focuses on a new method to
reduce uncertainties in multi-dimensional (gridded) multi-model projections of
the future climate with observations based on \ac{ML}. As an example, the
method is applied to \ac{GPP} at the end of the \nth{21} century, which is
already published in the \emph{Journal of Geophysical Research: Biogeosciences}
\autocite{Schlund2020}. Finally, \cref{ch:07:conclusion} provides a summary of
the results of this thesis and gives an outlook of possible future work.

>>>>

Climate change is one of the greatest challenges for humankind today. The
warming of the climate system is \enquote{unequivocal}, and \enquote{many of
  the observed changes are unprecedented over decades to millennia}
\autocite{IPCC2014}. The changing climate increases the \enquote{likelihood of
  severe, pervasive and irreversible impacts for people, species and
  ecosystems} with \enquote{mostly negative impacts for biodiversity, ecosystem
services and economic development}, and amplifies \enquote{risks for
livelihoods and for food and human security} \autocite{IPCC2014}. Potential
drivers for climate change are all natural and anthropogenic substances and
processes that may alter the Earth's energy budget. Human influence on the
climate system is clear: Ever-increasing emissions of \acp{GHG} since the end
of the pre-industrial era largely driven by economic and population growth
led to atmospheric concentrations of \ac{CO2}, \ac{CH4}, and \ac{N2O} that
are \enquote{unprecedented in at least the last 800000 years}
\autocite{IPCC2014}. The effects of these \ac{GHG} emissions and other
anthropogenic drivers have been \enquote{detected throughout the climate
system and are extremely likely to have been the dominant cause of the
observed warming since the mid-\nth{20} century} \autocite{IPCC2014}. The
greenhouse effect is based on the optical properties of the \acp{GHG}: While
the atmosphere is mostly transparent for the incoming solar (shortwave)
radiation, the outgoing infrared (longwave) radiation that is reflected from
the planet's surface is partly absorbed by the corresponding \ac{GHG}
molecules through excitation of their corresponding vibrational modes and
re-emitted isotropically. This process traps energy near the surface and
leads to a warming of the Earth's surface and the lower atmosphere. The
individual impact of the different drivers of climate change is measured with
the so-called \emph{radiative forcing}, which quantifies the change in energy
fluxes caused by changes in these drivers relative to pre-industrial
conditions \autocite{IPCC2013}. The total radiative forcing is positive
(which corresponds to a warming) and its largest contribution is caused by
the increase in the atmospheric concentration of \ac{CO2} due to
anthropogenic fossil fuel emissions since the mid-\nth{18} century
\autocite{IPCC2014}. Apart from their physical warming effect on the climate,
carbon-based \acp{GHG} like \ac{CO2} also directly influence the global
carbon cycle, an important biogeochemical cycle of the Earth. A crucial flux
of the carbon cycle is the \emph{\ac{GPP}}, which describes the carbon uptake
of the terrestrial biosphere due to photosynthesis. Since this land carbon
uptake absorbs about $30 \unit{\%}$ of the anthropogenic \ac{CO2} emissions
in today's climate \autocite{Friedlingstein2020}, this process substantially
slows down global warming and directly contributes to the magnitude of
climate sensitivity. Other important anthropogenic drivers of climate change
are the emission of aerosols, and land use/land cover changes. Apart from
anthropogenic drivers, there are also natural processes that impact the climate
system like changes in solar activity or volcanic eruptions. However, there is
clear evidence that these natural drivers alone cannot explain the observed
climate change \autocite{Haustein2017}.

@@

Simulations that extrapolate the state of the climate system into the future
are called \emph{climate model projections}. These include idealized
simulations with only a prescribed change in the atmospheric \ac{CO2}
concentration (\eg{} an instantaneous doubling of \ac{CO2} or \iac{CO2}
increase of $1 \unit{\%}$ per year) as well as more realistic projections that
consider different future scenarios (\eg{} a fossil fuel--based future or a
scenario that is based on sustainable development). Multi-model projections
from \ac{CMIP} show a large inter-model range in many variables that are
relevant for climate change \autocite{Collins2013, Flato2013}. A crucial and
policy-relevant example for this is climate sensitivity, which refers to the
change in the \ac{GSAT} that results from a change in the radiative forcing.
Common metrics for this are the \emph{\ac{ECS}}, which describes the
equilibrium response of the climate system after a doubling of the atmospheric
\ac{CO2} concentration, and the \emph{\ac{TCR}}, which describes the transient
response of the system to \iac{CO2} doubling. In \acs{AR}5, the assessed range
of both quantities is large with $\rangeunit{1.5}{4.5}{K}$ for \ac{ECS} and
$\rangeunit{1.0}{2.5}{K}$ for \ac{TCR} \autocite{Stocker2013}. The
corresponding inter-model ranges from the \acs{CMIP}5 models show similar
results \autocite{Flato2013}. For this reason, a careful statistical evaluation
and further refinement of the output of multi-model climate projections are
necessary to reduce associated uncertainties. A state-of-the-art technique for
this is the \emph{emergent constraints} method, which uses a physically based
inter-model relationship between an observable quantity of the Earth system and
a target variable to reduce uncertainties in the target variable with
observations \autocite{Allen2002}. An alternative approach is the weighting of
climate models based on their performance (\ie{} the distance of one model to
observational products) and interdependence (\ie{} the distance of one model to
other climate models) \autocite{Knutti2017a}. These techniques form the
baseline for the new analyses and results presented in this thesis, which
partly utilize methods from a new emerging research field in climate sciences:
\ac{AI} and \ac{ML}.

@@

This thesis aims to reduce to reduce uncertainties in multi-model climate
projections with observations by addressing the following three
\namecrefs{enum:01:question_1}:

@@

Parts of this thesis are already published in multiple peer-reviewed
publications (two first-author studies and six co-author studies). A complete
list of these publications is given on \cpageref{ch:bib:own}. Wherever material
from these studies is presented in this thesis, the pronoun \enquote{we} is
used to increase readability by avoiding the passive voice and to acknowledge
all involved contributors. However, unless stated otherwise, all content from
these publications (text, figures, and tables) shown in this thesis originates
from the author of this thesis. A detailed list of contributions to these
studies is given in the corresponding \namecrefs{ch:03:esmvaltool}.

This thesis is structured as follows: \Cref{ch:02:scientific_background}
introduces the scientific background. This includes relevant literature that is
used as a baseline for this thesis. \Cref{ch:03:esmvaltool} gives an overview
of the contributions made to the \ac{ESMValTool}, an open-source software
package for the analysis of \acp{ESM}. These contributions helped to improve
the routine evaluation of \acp{ESM} which is useful for the entire scientific
community and led to co-authorship in four peer-reviewed studies
\autocite{Eyring2020, Lauer2020, Righi2020, Weigel2020}.
\Cref{ch:04:papers_ecs_tcr_assessment} covers the assessment of climate
sensitivity metrics like the \ac{ECS} or \ac{TCR} in the latest generation of
\acp{ESM} from \acs{CMIP}6. This work is already published in two scientific
publications \autocite{Bock2020, Meehl2020}. Since the \ac{ECS} and \ac{TCR}
are considerably higher in this new climate model generation,
\cref{ch:05:paper_ecs} describes the assessment of emergent constraints on the
\ac{ECS} for these \acs{CMIP}6 models and compares these to results derived
from \acs{CMIP}5 models. The content of this \namecref{ch:05:paper_ecs} is
published in \emph{Earth System Dynamics} \autocite{Schlund2020a}.
\Cref{ch:06:paper_gpp} focuses on a new method to reduce uncertainties in
multi-dimensional (gridded) multi-model projections of the future climate with
observations based on \ac{ML}. As an example, the method is applied to \ac{GPP}
at the end of the \nth{21} century, which is already published in the
\emph{Journal of Geophysical Research: Biogeosciences} \autocite{Schlund2020}.
Finally, \cref{ch:07:conclusion} provides a summary of the results of this
thesis and gives an outlook of possible future work.

### 2021-04-08 ch02_scientific_background/scientific_background.tex
Spell- and grammer-checked chapter 1

An important policy-relevant metric of the Earth system that can be assessed
with numerical climate model simulations is the climate sensitivity. Climate
sensitivity refers to the change in \ac{GSAT} that results from a change in the
radiative forcing. In other words, it describes how sensitive the climate
system is to an external forcing. The source of this forcing might either be
natural (changes in the solar activity, volcanic eruptions, \etc{}) or
anthropogenic (emissions of \acp{GHG}, land use changes, \etc{}). Thus, the
assessment of climate sensitivity is essential for a precise quantification of
the human-made climate change to determine optimal mitigation and adaptation
strategies.

@@

For some domains of the Earth system, feedbacks can be positive and/or
negative, since a variety of different mechanisms are involved. An example for
this is the cloud feedback. Changes in the clouds induced by climate change can
cause changes in their longwave (greenhouse warming) and shortwave (reflective
cooling) effects on the Earth's radiation budget, which both need to be
considered for the overall cloud feedback \autocite{Boucher2013}. Relevant
cloud properties that may change as a response to an external forcing and that
may alter the Earth's radiative budget are cloud cover, cloud optical
thickness, cloud top and cloud base height, vertical extent, and the
geographical distribution of clouds. Examples for robust cloud feedback
processes are the increase in cloud top height of high-level clouds in a
warming climate which traps longwave radiation and enhances global warming, and
the reduction in mid- and low-level cloud cover which diminishes the reflection
of incoming solar radiation and also increases the surface warming
\autocite{Boucher2013}. In global climate model ensembles, the overall cloud
feedback shows a large range with positive and negative values, but tends to be
slightly positive on average \autocite{Soden2006, Dufresne2008, Vial2013,
  Zelinka2020}. This large uncertainty in the cloud feedback is a major reason
for uncertainties in the climate sensitivity of climate models
\autocite{Boucher2013, Flato2013}.

@@

For this reason, the equilibrium climate sensitivity is commonly approximated
with the \emph{\acf{ECS}}, which can be derived from only 150 model years of a
simulation with an abrupt quadrupling of the atmospheric \ac{CO2} concentration
(\nxcotwo{4}) \autocite{Gregory2004}. The basis of the definition of \ac{ECS}
is the simple energy balance model introduced in
\cref{subsec:02:mathematical_framework_feedbacks}. Assuming radiative
equilibrium ($N = 0$), \cref{eq:02:N,eq:02:linear_response} imply
\begin{equation}
  \Delta T = -\frac{F}{\lambda}.
  \label{eq:02:delta_t_eq}
\end{equation} Thus, the change in \ac{GSAT} in radiative equilibrium can be
easily calculated from the external forcing $F$ and the climate feedback
parameter $\lambda$. The steady state values for $F$ and $\lambda$ can be
estimated from a \nxcotwo{4} simulation that is not in equilibrium by
extrapolation with a linear regression \autocite{Gregory2004}. For this
so-called \emph{Gregory regression} (see \cref{fig:02:gregory_regression}), the
global and annual mean net \ac{TOA} radiation $N$ versus the change in the
annual mean \ac{GSAT} $\Delta T$ for all 150 years of the \nxcotwo{4} run are
plotted. To account for energy leakage and remove any model drift present in
the control climate, a linear fit of the corresponding pre-industrial control
run is subtracted from the \nxcotwo{4} simulation beforehand
\autocite{Andrews2012}. Since the combination of
\cref{eq:02:N,eq:02:linear_response} yields
\begin{equation}
  N = F + \lambda \cdot \Delta T,
  \label{eq:02:N_vs_delta_t}
\end{equation}
$F$ is now given by the \yintercept{} of this linear regression
($F_\mathtxt{4x}$) and $\lambda$ by its slope. Thus, the \ac{ECS} is given by
\begin{equation}
  \mathtxt{ECS} = -\frac{F_\mathtxt{4x}}{2 \lambda}.
  \label{eq:02:ecs}
\end{equation}
The factor of $2$ in the denominator accounts for the fact the traditional
equilibrium climate sensitivity is defined for an abrupt \ac{CO2} doubling,
whereas here a simulation with an abrupt quadrupling is considered.

@@

A further metric describing the climate sensitivity is the \emph{\ac{TCR}}. In
contrast to the \ac{ECS}, this metric does not assume radiative equilibrium of
the Earth system but describes the transient response of an evolving climate.
Following \textcite{Bindoff2013}, \ac{TCR} is defined as the change in the
\ac{GSAT} at the time of \ac{CO2} doubling in a simulation with a $1 \unit{\%}$
per year increase of the atmospheric \ac{CO2} concentration (\onepctcotwo{}).
For this, the annual mean \acp{GSAT} are averaged over a 20-year period
centered at the time of the \ac{CO2} doubling (years \range{61}{80} when the
first year is indexed with $1$). To account for model drift, the annual mean
changes in \ac{GSAT} are calculated relative to a corresponding pre-industrial
control simulation smoothed with a linear fit that considers 140 model years
(length of the \onepctcotwo{} simulation). An illustration of that calculation
is shown in \cref{fig:02:tcr}.

@@

In the last two decades, many emergent constraints on various aspects of the
Earth system have been published. Early studies tackled the hydrological cycle
\autocite{Allen2002} and the snow-albedo feedback \autocite{Hall2006}. Over the
years, the climate sensitivity expressed by the \ac{ECS} has been a prominent
target variable. Since cloud feedbacks are a major source of uncertainty for
climate sensitivity, a variety of scientific papers focus on constraining
\ac{ECS} with cloud-related processes \autocite{Brient2015, Brient2016,
  Fasullo2012, Lipat2017, Sherwood2014, Su2014, Tian2015, Volodin2008, Qu2013,
  Zhai2015}, which are discussed in detail in
\cref{sec:05:comparison_of_emergent_constraints}. More recent studies aim to
constrain \ac{ECS} with the historical temperature variability
\autocite{Cox2018} or the historical warming trend
\autocite{JimenezdelaCuesta2019, Nijsse2020, Tokarska2020}. Emergent
constraints are not only limited to physical processes, but can also be applied
to other domains, like the global carbon cycle \autocite{Cox2013,
  Kwiatkowski2017, Wenzel2014, Wenzel2016, Winkler2019}. An extensive
discussion on the emergent constraint by \textcite{Wenzel2016}, which focuses
on the concentration-carbon feedback, is given in \cref{subsec:06:step_1}.

>>>>

An important policy-relevant metric of the Earth system that can be assessed
with numerical climate model simulations is climate sensitivity. Climate
sensitivity refers to the change in \ac{GSAT} that results from a change in the
radiative forcing. In other words, it describes how sensitive the climate
system is to an external forcing. The source of this forcing might either be
natural (changes in the solar activity, volcanic eruptions, \etc{}) or
anthropogenic (emissions of \acp{GHG}, land use changes, \etc{}). Thus, the
assessment of climate sensitivity is essential for a precise quantification of
the human-made climate change to determine optimal mitigation and adaptation
strategies.

@@

For some domains of the Earth system, feedbacks can be positive and/or
negative, since a variety of different mechanisms are involved. An example for
this is the cloud feedback. Changes in the clouds induced by climate change can
cause changes in their longwave (greenhouse warming) and shortwave (reflective
cooling) effects on the Earth's radiation budget, which both need to be
considered for the overall cloud feedback \autocite{Boucher2013}. Relevant
cloud properties that may change as a response to an external forcing and that
may alter the Earth's radiative budget are cloud cover, cloud optical
thickness, cloud top and cloud base height, vertical extent, and the
geographical distribution of clouds. Examples for robust cloud feedback
processes are the increase in cloud top height of high-level clouds in a
warming climate which traps longwave radiation and enhances global warming, and
the reduction in mid- and low-level cloud cover which diminishes the reflection
of incoming solar radiation and also increases the surface warming
\autocite{Boucher2013}. In global climate model ensembles, the overall cloud
feedback shows a large range with positive and negative values, but tends to be
slightly positive on average \autocite{Soden2006, Dufresne2008, Vial2013,
  Zelinka2020}. This large uncertainty in the cloud feedback is a major reason
for uncertainties in climate sensitivity of climate models
\autocite{Boucher2013, Flato2013}.

@@

For this reason, the equilibrium climate sensitivity is commonly approximated
with the \emph{\acf{ECS}}, which can be derived from only 150 model years of a
simulation with an abrupt quadrupling of the atmospheric \ac{CO2} concentration
(\nxcotwo{4}) \autocite{Gregory2004}. The basis of the definition of \ac{ECS}
is the simple energy balance model introduced in
\cref{subsec:02:mathematical_framework_feedbacks}. Assuming radiative
equilibrium ($N = 0$), \cref{eq:02:N,eq:02:linear_response} imply
\begin{equation}
  \Delta T = -\frac{F}{\lambda}.
  \label{eq:02:delta_t_eq}
\end{equation} Thus, the change in \ac{GSAT} in radiative equilibrium can be
easily calculated from the external forcing $F$ and the climate feedback
parameter $\lambda$. The steady state values for $F$ and $\lambda$ can be
estimated from a \nxcotwo{4} simulation that is not in equilibrium by
extrapolation with a linear regression \autocite{Gregory2004}. For this
so-called \emph{Gregory regression} (see \cref{fig:02:gregory_regression}), the
global and annual mean net \ac{TOA} radiation $N$ versus the change in the
annual mean \ac{GSAT} $\Delta T$ for all 150 years of the \nxcotwo{4} run are
plotted. To account for energy leakage and remove any model drift present in
the control climate, a linear fit of the corresponding pre-industrial control
run is subtracted from the \nxcotwo{4} simulation beforehand
\autocite{Andrews2012}. Since the combination of
\cref{eq:02:N,eq:02:linear_response} yields
\begin{equation}
  N = F + \lambda \cdot \Delta T,
  \label{eq:02:N_vs_delta_t}
\end{equation}
$F$ is now given by the \yintercept{} of this linear regression
($F_\mathtxt{4x}$) and $\lambda$ by its slope. Thus, the \ac{ECS} is given by
\begin{equation}
  \mathtxt{ECS} = -\frac{F_\mathtxt{4x}}{2 \lambda}.
  \label{eq:02:ecs}
\end{equation}
The factor of $2$ in the denominator accounts for the fact that the traditional
equilibrium climate sensitivity is defined for an abrupt \ac{CO2} doubling,
whereas here a simulation with an abrupt quadrupling is considered.

@@

A further metric describing climate sensitivity is the \emph{\ac{TCR}}. In
contrast to the \ac{ECS}, this metric does not assume radiative equilibrium of
the Earth system but describes the transient response of an evolving climate.
Following \textcite{Bindoff2013}, \ac{TCR} is defined as the change in the
\ac{GSAT} at the time of \ac{CO2} doubling in a simulation with a $1 \unit{\%}$
per year increase of the atmospheric \ac{CO2} concentration (\onepctcotwo{}).
For this, the annual mean \acp{GSAT} are averaged over a 20-year period
centered at the time of the \ac{CO2} doubling (years \range{61}{80} when the
first year is indexed with $1$). To account for model drift, the annual mean
changes in \ac{GSAT} are calculated relative to a corresponding pre-industrial
control simulation smoothed with a linear fit that considers 140 model years
(length of the \onepctcotwo{} simulation). An illustration of that calculation
is shown in \cref{fig:02:tcr}.

@@

In the last two decades, many emergent constraints on various aspects of the
Earth system have been published. Early studies tackled the hydrological cycle
\autocite{Allen2002} and the snow-albedo feedback \autocite{Hall2006}. Over the
years, climate sensitivity expressed by the \ac{ECS} has been a prominent
target variable. Since cloud feedbacks are a major source of uncertainty for
climate sensitivity, a variety of scientific papers focus on constraining
\ac{ECS} with cloud-related processes \autocite{Brient2015, Brient2016,
  Fasullo2012, Lipat2017, Sherwood2014, Su2014, Tian2015, Volodin2008, Qu2013,
  Zhai2015}, which are discussed in detail in
\cref{sec:05:comparison_of_emergent_constraints}. More recent studies aim to
constrain \ac{ECS} with the historical temperature variability
\autocite{Cox2018} or the historical warming trend
\autocite{JimenezdelaCuesta2019, Nijsse2020, Tokarska2020}. Emergent
constraints are not only limited to physical processes, but can also be applied
to other domains, like the global carbon cycle \autocite{Cox2013,
  Kwiatkowski2017, Wenzel2014, Wenzel2016, Winkler2019}. An extensive
discussion on the emergent constraint by \textcite{Wenzel2016}, which focuses
on the concentration-carbon feedback, is given in \cref{subsec:06:step_1}.
